Laziness in R

### 2018-10-08 report/report.Rmd
Add report with function-viewer application

---
title: "Analyses of Promises in R"
author: "Aviral Goel"
output: 
  html_document:
    theme: flatly
    highlight: tango
    keep_md: true
    toc: true
    toc_float: true
    number_sections: true
params:
  analysis_output_dir: TRUE
---

# Lifespan

```{r set-options, echo=FALSE, cache=FALSE}
options(width = 1200)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(methods) ## including this prevents a warning - "executing load actions ..."
library(DT)
library(tidyverse)
library(knitr)
library(png)
library(grid)
library(gridExtra)
```

```{r echo=FALSE, message=FALSE}
# https://www.rdocumentation.org/packages/knitr/versions/1.17/topics/include_graphics

get_graph <- function(analysis_name, graph_filename) {
    file.path(params$analysis_output_dir, analysis_name, "graph", graph_filename)
}

show_graph <- function(analysis_name, graph_filename) {
    get_graph(analysis_name, graph_filename) %>%
    include_graphics()
}

get_table <- function(analysis_name, table_filename) {
    file.path(params$analysis_output_dir, analysis_name, "table", table_filename) %>%
    read_csv()
}

show_table <- function(analysis_name, table_filename) {
    get_table(analysis_name, table_filename) %>%
    datatable(rownames = FALSE)
}
```

```{r echo=FALSE, message=FALSE}
lifespan_summary <- get_table("promise-lifespan", "lifespan_summary.csv")
required_lifespan_summary <- get_table("promise-lifespan", "required_lifespan_summary.csv")
```

This section discusses the lifespan of promises in terms of the number of GC Cycles.
The lifespan of a promise is defined as the number of GC cycles for which the 
promise is alive.

## Promise Categories

We categorize the promises observed while tracing as follows:

1. **Foreign Promises** are created before the tracing starts. Since we don't see their creation, 
   we can't study their lifespan in terms of GC cycles. They are **excluded from the analysis**
2. **Immortal Promises** are created but not garbage collected while tracing. They are 
   **excluded from the analysis** because we can't compute their lifespan.
3. **Mortal Promises** are created and garbage collected while tracing.
   They are the only promises that are **included in the analysis**.
   They are further classified as follows:
   i. **Forced Promises** are the mortal promises that are looked up at least once.
   ii. **Unforced Promises** are mortal promises that are never looked up.

- **TOTAL PROMISE COUNT** = **FOREIGN PROMISE COUNT** + **IMMORTAL PROMISE COUNT** + **MORTAL PROMISE COUNT**

- **IMMORTAL PROMISE COUNT** = **UNFORCED PROMISE COUNT** + **FORCED PROMISE COUNT**

```{r echo=FALSE, message=FALSE, fig.width=10, fig.height=4.5}
show_graph("promise-lifespan", "category_counts.png")
```

### Summarized Data

```{r echo=FALSE, message=FALSE}
show_table("promise-lifespan", "count_summary.csv")
```

### Raw Data

```{r echo=FALSE, message=FALSE}
show_table("promise-lifespan", "counts.csv")
```

## Mean GC Cycles Distribution

Following graph shows the weighted mean of the number GC cycles 
for which promises are (should be) alive in the scripts.

```{r echo=FALSE, message=FALSE, out.width="500px"}
show_graph("promise-lifespan", "lifespan_mean_gc_cycles.png")
show_graph("promise-lifespan", "required_lifespan_mean_gc_cycles.png")
show_graph("promise-lifespan", "extra_lifespan_mean_gc_cycles.png")
```

## Median GC Cycles Distribution

Following graph shows the median number GC cycles for which 
promises are (should be) alive in the scripts.

```{r echo=FALSE, message=FALSE, out.width="500px"}
show_graph("promise-lifespan", "lifespan_median_gc_cycles.png")
show_graph("promise-lifespan", "required_lifespan_median_gc_cycles.png")
show_graph("promise-lifespan", "extra_lifespan_median_gc_cycles.png")
```

## Min GC Cycles Distribution

Following graph shows the minimum number GC cycles for which 
promises are (should be) alive in the scripts.

```{r echo=FALSE, message=FALSE, out.width="500px"}
show_graph("promise-lifespan", "lifespan_min_gc_cycles.png")
show_graph("promise-lifespan", "required_lifespan_min_gc_cycles.png")
show_graph("promise-lifespan", "extra_lifespan_min_gc_cycles.png")
```

## Max GC Cycles Distribution

Following graph shows the maximum number GC cycles for which 
promises are (should be) alive in the scripts.

```{r echo=FALSE, message=FALSE, out.width="500px"}
show_graph("promise-lifespan", "lifespan_max_gc_cycles.png")
show_graph("promise-lifespan", "required_lifespan_max_gc_cycles.png")
show_graph("promise-lifespan", "extra_lifespan_max_gc_cycles.png")
```

## Lifespan Data (Summarized)
```{r echo=FALSE, message=FALSE}
show_table("promise-lifespan", "lifespan_summary.csv")
```

## Required Lifespan Data (Summarized)
```{r echo=FALSE, message=FALSE}
show_table("promise-lifespan", "required_lifespan_summary.csv")
```

## Extra Lifespan Data (Summarized)
```{r echo=FALSE, message=FALSE}
show_table("promise-lifespan", "extra_lifespan_summary.csv")
```

## Lifespan Data (Raw)
```{r echo=FALSE, message=FALSE}
show_table("promise-lifespan", "lifespan_summary.csv")
```

## Required Lifespan Data (Raw)
```{r echo=FALSE, message=FALSE}
show_table("promise-lifespan", "required_lifespan.csv")
```

>>>>

---
title: "Laziness in R"
output:
  html_document:
    toc: true
    theme: united
params:
  output_directory: ../latest
  image_format: png
editor_options:
  chunk_output_type: console
---

<style type="text/css">
.main-container {
  max-width: 1200px;
  margin-left: auto;
  margin-right: auto;
  padding-left: 15px;
  padding-right: 15px;
}

.shiny-frame {
    height: 550px;
}
</style>

```{r echo = FALSE, message = FALSE}

library(DT)
library(fs)
library(png)
library(stringr)
library(dplyr)
library(magrittr)
library(rlang)
library(knitr)

```

```{r paths, echo = FALSE}

function_definitions_dirpath <- path(params$output_directory, "function-definitions")
parameter_usage_dirpath <- path(params$output_directory, "parameter-usage")
visualizations <- path(parameter_usage_dirpath, "visualizations")
summary <- path(parameter_usage_dirpath, "summary")

```

```{r datatsets, echo = FALSE}
import_data_table <- function(name) {
    write(path(summary, name), "log.txt")
    promisedyntracer::read_data_table(path(summary, name))
}

closure_force_order_distribution <-
     import_data_table("closure_force_order_distribution.csv") %>%
     mutate(force_order = str_replace_all(str_sub(force_order, 2), "\\|", " ~ "))

closure_force_order_distribution_summary <-
     import_data_table("closure_force_order_distribution_summary.csv")

closures <- import_data_table("closures.csv")

closure_call_count_summary_distribution <- 
    import_data_table("closure_call_count_summary_distribution.csv") %>%
    mutate(relative_function_count = round(relative_function_count * 100, 3))

closure_argument_count_summary_distribution <- 
    import_data_table("closure_argument_count_summary_distribution.csv") %>%
    mutate(relative_function_count = round(relative_function_count * 100, 3))

classification_summary <- import_data_table("classification_summary.csv")

parameter_mode_distribution_summary <-
    import_data_table("parameter_mode_distribution_summary.csv")

argument_mode_distribution_summary <-
    import_data_table("argument_mode_distribution_summary.csv")
#

# closure_classification_summary_distribution <- import_data_table("closure_classification_summary_distribution.csv")
# 
# position_classification_summary_distribution <- import_data_table("position_classification_summary_distribution.csv")
# 
# summary <- import_data_table("summary.csv")
#

parameter_force_order_summary <- 
     import_data_table("closure_force_order_distribution.csv") %>%
     mutate(force_order = str_replace_all(str_sub(force_order, 2), "\\|", " ~ "))

function_class_summary_distribution <-
    import_data_table("function_class_summary_distribution.csv") %>%
    mutate(relative_function_count = round(relative_function_count * 100, 3))

parameter_class_summary_distribution <-
    import_data_table("parameter_class_summary_distribution.csv") %>%
    mutate(relative_parameter_count = round(relative_parameter_count * 100, 3))

argument_class_distribution_summary <-
    import_data_table("argument_class_distribution_summary.csv")

argument_use_mode_distribution_summary <- 
     import_data_table("argument_use_mode_distribution_summary.csv")
```

```{r echo=FALSE}
count_labels <-
  Vectorize(function(x, digits = 2) {

    paste_round <- function(value, div, suffix, sep)
        paste(round(value/div, digits), suffix, sep = sep)

    if(is.na(x)) {
      "NA"
    } else if(x < 10^3) {
      paste0(x)
    } else if(x < 10^6) {
      paste_round(x, 1000, "K", sep=" ")
    } else if(x < 10^9) {
      paste_round(x, 10^6, "M", sep=" ")
    } else {
      paste_round(x, 10^9, "B", sep=" ")
    }
  },
  "x")
```

This report summarizes the findings on the use of laziness in R.

Data from 23 packages and 41 vignettes.

* Number of functions: `r count_labels(length(unique(closures$function_id)))` 
* Number of calls: `r count_labels(sum(closures$call_count))`
* Number of parameters: `r count_labels(sum(closures$parameter_count))`
* Number of arguments: `r count_labels(sum(closures$parameter_count * closures$call_count))`

## Closures

The table below gives the relevant information for all these functions.

```{r closure_data, echo = FALSE}
datatable(closure_force_order_distribution) #, width="1000px")
```

## Closure call count distribution

The following figures show the distribution of the number of calls made to closures.

```{r echo = FALSE, out.width = "1000px"}
include_graphics(path(visualizations, "closure_call_distribution.png"))
```

```{r echo = FALSE}
datatable(closure_call_count_summary_distribution)
```

### Observations

* Very few functions, **`r closure_call_count_summary_distribution %>% filter(call_count == 1) %>% pull(relative_function_count) %>% first()`%** , are called only once. We can remove these functions from the remaining analyses without affecting the representativity of our results.

* Majority of functions, **`r closure_call_count_summary_distribution %>% filter(!(call_count %in% c("1", "> 10"))) %>% pull(relative_function_count) %>% sum()`%**, are called more than once.

* **`r closure_call_count_summary_distribution %>% filter(call_count == "> 10") %>% pull(relative_function_count) %>% sum()`%** functions are called more than 10 times.

## Closure parameter count distribution

The following figures show the distribution of the number of parameters of closures.

```{r echo = FALSE, out.width = "1000px"}
include_graphics(path(visualizations, "closure_parameter_count_distribution.png"))
```

```{r echo = FALSE}
datatable(closure_argument_count_summary_distribution)
```

### Observations
- Majority of functions, **`r closure_argument_count_summary_distribution %>% filter(parameter_count %in% c(1, 2, 3)) %>% pull(relative_function_count) %>% sum()`%**, have one, two or three parameters.
- A very small number of functions, **`r closure_argument_count_summary_distribution %>% filter(parameter_count == 0) %>% pull(relative_function_count) %>% first()`%**, have zero parameters.

## Closure force order distribution

```{r echo = FALSE, out.width = "1000px"}
include_graphics(path(visualizations, "closure_force_order_distribution.png"))
```

```{r echo = FALSE}
datatable(closure_force_order_distribution_summary)
```

<!--## Parameter Mode

```{r echo = FALSE, out.width = "1000px"}
include_graphics(path(visualizations, "parameter_mode_distribution.png"))
```

```{r echo = FALSE}
datatable(parameter_mode_distribution_summary)
```
-->
## Argument Mode

```{r echo = FALSE, out.width = "1000px"}
include_graphics(path(visualizations, "argument_mode_distribution.png"))
```

```{r echo = FALSE}
datatable(argument_mode_distribution_summary)
```

## Argument Class

```{r echo = FALSE, out.width = "1000px"}
include_graphics(path(visualizations, "argument_class_distribution.png"))
```

```{r echo = FALSE}
datatable(argument_class_distribution_summary)
```

## Argument Use Mode

```{r echo = FALSE, out.width = "1000px"}
include_graphics(path(visualizations, "argument_use_mode_distribution.png"))
```

```{r echo = FALSE}
datatable(argument_use_mode_distribution_summary)
```

## Parameter Class

This section discusses function parameter position usage. The parameter of a function in R can be:
1. looked up for its value
2. looked up for the argument expression
3. not looked up

The graph below shows how the parameter positions are distributed across these categories in all calls to functions.

```{r echo=FALSE}
# datatable(classification_summary)
```

```{r echo = FALSE, out.width = "1000px"}
include_graphics(path(visualizations, "parameter_class_distribution.png"))
```

```{r echo = FALSE}
datatable(parameter_class_summary_distribution)
```

## Argument Class

```{r echo = FALSE, out.width = "1000px"}
include_graphics(path(visualizations, "argument_count_by_mode_class_and_use.png"))
```

```{r echo = FALSE}
datatable(argument_use_mode_distribution_summary)
```

## Function Class

```{r echo = FALSE, out.width = "1000px"}
include_graphics(path(visualizations, "function_class_distribution.png"))
```

```{r echo = FALSE, out.width = "1000px"}
include_graphics(path(visualizations, "function_class_call_distribution.png"))
```

```{r echo = FALSE}
datatable(function_class_summary_distribution)
```
